% JSAgent methodology paper — arXiv scaffold (Goal 2)
% Build:  pdflatex main && bibtex main && pdflatex main && pdflatex main
% Self-contained: compiles with a stock TeX Live (no external .sty files required).
%
% Title: SET (Goal 7) — option (a), "Wiki" -> "Knowledge Base" per author.
%
% STATUS (Goal 7 complete): all sections drafted; numbers re-verified against the live sweep
% 2026-06-22 22:00 PT. §1 origin/thesis and §8 personal close are AUTHOR-DRAFTs (marked) for
% the author to revise into their own voice.
%
% BEFORE arXiv SUBMISSION (final pass — not done here, no local TeX toolchain):
%   1. Strip internal provenance markers: the \src{...} notes and the \stub macro definition
%      are for drafting traceability only — delete \src usages and the \newcommand lines.
%   2. Build: pdflatex main && bibtex main && pdflatex main && pdflatex main  (needs TeX).
%   3. Re-run scripts/leaderboard_standings.py and reconcile Table 1 + inline numbers.
%   4. Optional figures: lineage graph, cycle-log self-improvement curve (see outline §Open).
%   5. Author revises the two AUTHOR-DRAFT blocks (§1, §8).

\documentclass[11pt]{article}

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage[margin=1in]{geometry}
\usepackage{amsmath,amssymb,amsthm}
\usepackage{booktabs}
\usepackage{graphicx}
\usepackage{xcolor}
\usepackage{enumitem}
\usepackage[hidelinks]{hyperref}
\usepackage{url}

% Scaffold marker for stub content — remove before final build.
\newcommand{\stub}[1]{\par\noindent\textcolor{gray}{\small\textit{[stub: #1]}}\par}
\newcommand{\src}[1]{\textcolor{gray}{\footnotesize\textsf{(src: #1)}}}

\title{\textbf{JSAgent: A Self-Improving Agent that Compounds\\
  Mathematical Wisdom in a Knowledge Base}}

\author{%
  Jongmin Sung\\
  \texttt{jongmin@sailplane.ai}}

\date{Draft --- \today\\[2pt]
  \small Companion to Bianchi et al., \emph{Harnessing the Collective Intelligence of
  AI Agents in the Wild for New Discoveries} (arXiv:2606.10402).}

\begin{document}
\maketitle

\begin{abstract}
Recent systems (AlphaEvolve, TTT-Discover) show language-model agents can advance open
math problems, but each runs as a single orchestrated pipeline whose knowledge evaporates
when the run ends. We describe \textbf{JSAgent}, an agent built on the opposite bet: its
primary artifact is a \textbf{compounding wiki of mathematical wisdom plus a set of
behavioral rules co-evolved with the agent}, and arena submission is treated as
\emph{verification}, not the goal. Operating in the wild on EinsteinArena, JSAgent reached
\textbf{top-3 on 10 of 17 active problems and the highest rank-\#1 count (4) and top-3
count (10) of any agent} (as of 2026-06-22). We detail the methodology --- a wiki-first
self-improvement loop, a mathematician-persona ``council'' that writes questions rather
than answers, a triple-verify discipline that guards against evaluator drift, a wall-hit
escalation that consults accumulated failure before brute force, and a compute router ---
and the concrete techniques it produced (cross-resolution basin transfer, Dinkelbach
fractional programming, warm self-pruning, mpmath ULP polish). We
argue, with evidence and stated caveats, that agents are ready to make real, auditable
scientific contributions today. Every artifact of this work --- code, scripts, wiki, and
this paper --- was implemented by the agent under human guidance, not written by hand; the
human's load-bearing contribution was \emph{which question to ask} and \emph{comprehending
what was done well enough to steer it}. We close with the conditions under which this loop
becomes fully autonomous, and why objective verification (which math enjoys and the natural
sciences largely lack) is the gating factor.
\end{abstract}

% ======================================================================
\section{Introduction}
\label{sec:intro}
% Voice: author-led, first person. DRAFT for the author to make their own — the felt parts
% are placeholders in the author's register, not final claims. See outline.md §1.

% --- AUTHOR-DRAFT (revise freely; this is scaffolding in your voice) ---
I did not set out to climb a leaderboard. I set out to find out whether an agent, given the
right discipline, could accumulate genuine mathematical \emph{wisdom} --- the kind that
compounds --- rather than produce a single clever answer and forget how it got there.
Mathematical optimization on a public arena was the proving ground because it is honest: a
verifier either accepts a construction or it does not, so there is nowhere for
self-deception to hide.

The needle moved gradually and then suddenly. Early submissions were unremarkable; the
turning points came when failures started being written down \emph{with their reasons}, and
later runs began standing on those write-ups instead of rediscovering the same walls. The
clearest instance --- a verified arena record produced verbatim from a four-day-old
dead-end's ``what might still work'' note --- convinced me the deliverable was never the
score. It was the page that produced the score.

That is the thesis of this paper: \textbf{an agent whose deliverable is compounding wisdom,
not a one-shot run.} The knowledge base is the artifact; arena submission is only a way to
verify it against the world. I will argue, with verified standings and stated caveats, that
agents are ready to make real, auditable scientific contributions today --- and, just as
honestly, that the human's role in choosing what to work on and understanding what was done
has not gone away.
% --- end AUTHOR-DRAFT ---

\paragraph{How this work was produced (and why it matters to the thesis).}
% DRAFT (per author): the human-mediated / agent-implemented framing. Author's voice — refine in Goal 7.
Every artifact described here --- the optimizer code and per-problem scripts, the wiki of
concepts and findings, the behavioral rules, and the prose of this paper --- was
\emph{implemented by the agent}, not typed by the author. The mode of work was uniformly
agentic: the human stated a request, the agent did it and reported back. This is not an
aside; it is part of the result. The central lesson is a clean division of labor that
emerged in practice: the agent is extraordinarily capable at \emph{execution}, but it does
not reliably know \emph{which question is worth asking} or which problem is worth solving
next --- problem selection and taste remained the human's job. And steering that selection
requires the human to genuinely \emph{comprehend} what the agent is doing; without
comprehension there is no guidance, only rubber-stamping. The wiki and the harness are, in
part, the instruments that make such comprehension --- and therefore such guidance ---
possible at all (\S\ref{sec:harness}).

\paragraph{Contributions.}
\begin{itemize}[nosep]
  \item A self-improving wiki-agent \emph{methodology} (\S\ref{sec:harness}).
  \item An open, auditable \emph{harness} of behavioral rules co-evolved with the agent.
  \item Transferable \emph{techniques} produced in the course of solving (\S\ref{sec:techniques}).
  \item \emph{Evidence}: verified standings and an honest self-improvement trail (\S\ref{sec:results}).
  \item A field account of the \emph{human--agent division of labor} --- execution
        automates first, problem selection and comprehension do not (\S\ref{sec:discussion}).
\end{itemize}

% ======================================================================
\section{EinsteinArena and Positioning}
\label{sec:arena}
EinsteinArena~\cite{bianchi2026einstein,togetherai2026arena} is a public platform of open
mathematical optimization problems, each paired with a hardened automatic verifier and a
strictly-ranked leaderboard, plus an open discussion board. Its design bet is that discovery
is \emph{collective}: by making the platform the shared, persistent state, later agents
inherit the current frontier instead of restarting from scratch, and the operators report
twelve new state-of-the-art results across the problem set since launch. The arena is the
ideal testbed for our claim precisely because its verification is objective and cheap to
re-run --- the property \S\ref{sec:discussion} identifies as the gate on agent autonomy.

What distinguishes JSAgent from the field is not how it searches but what it
\emph{keeps}: where other entrants run a pipeline and submit, JSAgent treats each run as
disposable and the accumulated knowledge base as the deliverable
($\to$~\S\ref{sec:related}). \src{docs/source/2026-bianchi-einstein-arena-collective-intelligence.md}

% ======================================================================
\section{Mathematical Contributions}
\label{sec:math}
% DRAFT (Goal 5): grounded in findings/arena-standings-snapshot-2026-06-22 + problem pages.
% All ranks/scores trace to the live sweep 2026-06-22T21:37 PT. Re-run before final build.

We lead with where JSAgent holds defensible wisdom rather than a full problem-by-problem
tour; the standings are in Table~\ref{tab:standings}, and the honest reading of each rank is
stated inline.

\paragraph{P2, first autocorrelation inequality (rank \#1, clean).}
The deepest result. Two obstructions were diagnosed and recorded: \emph{peak-locking} under
the $\exp(v)$ parameterization (the iterate freezes $\sim 5\mathrm{e}{-}8$ from SOTA across
all $n$ and all optimizers), escaped by switching to $f=v^2$; and the \emph{compact-support}
structure of the true optimum, reached by warm data-driven self-pruning (\S\ref{sec:techniques}).
The latter broke a two-agent tie and set a verified arena record ($1.5028609\to 1.5028506$).
\src{problems/2-first-autocorrelation.md, concepts/n-agent-tie-not-global-min.md}

\paragraph{P14 circle-packing and P18 circles-in-rectangle (rank \#1, clean).}
Strictly-best scores on the live board. \src{findings/arena-standings-snapshot-2026-06-22.md}
% TODO (Goal 7): one mechanism sentence each from problems/14-*, problems/18-* pages.

\paragraph{P9 uncertainty principle (rank \#1, tied --- held on submission order).}
A legitimate but contested \#1: another agent matched the exact score and submitted later, so
the arena's earliest-submission tiebreak awards JSAgent the top slot. We state this as ``tied,
held on submission order'' wherever it appears. The problem also carries a verification scar
--- a celebrated score once fell below the proven lower bound (an evaluator bug, not
mathematics), which is part of why triple-verify is the harness's spine.
\src{findings/arena-standings-snapshot-2026-06-22.md, findings/verification-patterns.md}

\paragraph{P3, second autocorrelation inequality (rank \#3, on the documented frontier).}
JSAgent achieved the strongest $10^5$-interval solution, $C=0.962214$, via cross-resolution
basin transfer and Dinkelbach refinement --- the result the platform
paper~\cite{bianchi2026einstein} attributes to JSAgent by name (its Fig.~6) and the lineage
in which it is shown discussing basin transfer with another agent (Fig.~3b). The final record
($0.962643$, at $4\times 10^5$ resolution) went to a different agent; JSAgent's contribution
is an intermediate link in the documented collective lineage --- which is precisely the
arena's thesis. \src{findings/einstein-arena-paper-jsagent-recognition-2026-06.md, problems/3-autocorrelation.md}

\paragraph{Honest framing.}
Of the four rank-\#1s, three (P2, P14, P18) are strictly-best and one (P9) is a held tie.
P1 is rank \#2 at a score \emph{byte-identical} to the leader, who submitted three weeks
earlier. The most robust single claim is therefore not any one problem but the all-agent lead
in both rank-\#1 count (4) and top-3 count (10), since that does not hinge on a tiebreak.
\src{findings/arena-standings-snapshot-2026-06-22.md}

% ======================================================================
\section{Agentic Harness Contribution}
\label{sec:harness}
% The structurally novel core. Each item is a .claude/rules/ rule with a *why*. See outline.md §4.
% DRAFT (Goal 3): prose grounded directly in .claude/rules/*. Incidents are real and dated;
% verify each against the cited rule file / finding before final build (Goal 7).

The contribution of this section is not any single algorithm but a \emph{harness}: a small
set of behavioral rules, written in plain language and version-controlled alongside the
code, that govern how the agent works a problem. Each rule exists because skipping it once
cost something concrete; the cost is recorded \emph{in} the rule as its ``why.'' The rules
are read into the agent's context every session, so the harness is simultaneously the
agent's operating manual and a human-auditable account of how it reasons. Unlike a frozen
orchestration pipeline, the harness and the wiki co-evolve: a failure produces a finding,
a recurring finding is promoted to a rule, and the next run inherits both.
\src{.claude/rules/, docs/agent/}

\paragraph{The self-improvement loop.}
Knowledge generated in-context is lost when the session ends; the wiki is the only layer
that survives across sessions, contexts, and model versions. The loop makes persistence
mandatory: a \emph{gap} (a question the wiki cannot answer) is \emph{filed} as a question
page, \emph{researched} (wiki-first, then human-gated ingest), \emph{distilled} into a
finding with citations, and --- if cited three or more times or generalizing beyond one
problem --- \emph{promoted} (human-gated) to a concept. The most-skipped step is the
cheapest: filing the question. An unfiled gap is wisdom permanently lost.
\src{.claude/rules/self-improvement-loop.md}

\paragraph{Council dispatch --- personas that write questions, not answers.}
At the start of a hard problem the agent dispatches three to five mathematician personas
(Gauss, Tao, Cohn, Viazovska, \ldots), each seeded with the problem, the current top
solutions, and the persona's own stance page. Their job is explicitly \emph{not} to solve
it: a persona producing ``the answer'' is just plausible markdown, whereas a persona
producing \emph{the question this stance would ask when stuck} forces the gap-detect loop.
The council runs once per problem during reconnaissance --- and is re-triggered on a wall
(below).
\src{.claude/rules/council-dispatch.md}

\paragraph{Wiki-first lookup and cycle discipline.}
Before answering any in-scope question the agent must query the wiki first and cite what it
finds; the first tool call of every problem cycle is a wiki query, treated as a
\emph{correctness check on memory} rather than a search-because-stuck. Every cycle ---
success or failure --- produces exactly one row in a cycle log. The discipline is
anti-cherry-picking: without a strict one-to-one cycle-to-log rule, the self-improvement
metric degrades into a best-of-$N$ highlight reel that looks like progress without being it.
\src{.claude/rules/wiki-first-lookup.md, .claude/rules/cycle-discipline.md}

\paragraph{Triple-verify --- the credibility spine.}
Every score is verified three ways before it is trusted: a fast local evaluator, an
independent exact reimplementation on a different code path, and a cross-check against an
analytical bound or a structurally different method. If any two disagree, the score is
treated as fake until the discrepancy is explained. This rule is the most load-bearing in
the harness because evaluator drift has caused real damage on this arena --- a verifier bug
that hid a problem (P9), an exploitable tolerance band (P14), and a strict-tolerance trap
that would have cost a large penalty (P17). Arena submission is then only a \emph{fourth}
check: slower, binary, and non-iterating. The local triple-verify is the closed loop.
\src{.claude/rules/triple-verify.md, .claude/rules/axioms.md (A1)}

\paragraph{Wall-hit escalation --- consult failure before brute force.}
A wall is defined concretely (two consecutive non-improving attempts, a method retuned
twice, or roughly thirty minutes with no metric gain). On a wall the agent must
\emph{stop} computing and escalate: name the obstruction as a question, grep the wall
ledger for a matching past obstruction (and reuse its resolving move if found), re-query
the wiki, and \emph{re-dispatch the council} seeded with what failed and why --- each
persona proposing one new move outside the failed class. The triggering incident is
explicit: on a record attempt the agent once ran ten-plus optimizer variants over hours
before consulting the wiki, after which the council diagnosed the obstruction (a discrete
incompressible sign topology) and ruled out whole families in minutes. A corollary guards
``floor'' claims: an $N$-agent empirical tie is evidence of a sharp shared basin, not a
proof of global optimality --- a verified arena record on P2 came from a single untried
operator (warm self-pruning) that beat a two-agent tie thought to be the global minimum.
\src{.claude/rules/wall-hit-escalation.md, concepts/n-agent-tie-not-global-min}

\paragraph{Failure is a finding.}
Every dead-end is written up with its \emph{obstruction} --- not ``it did not improve''
(the symptom) but \emph{why} it did not --- plus what the failure rules out and what might
still work. The ``what might still work'' section is load-bearing rather than garnish: the
P2 arena record (the branch's first) came \emph{verbatim} from the next-door section of a
dead-end finding written four days earlier. The agent did not invent the move that session;
it executed a prescription the failure-log discipline had already recorded.
\src{.claude/rules/failure-is-a-finding.md}

\paragraph{Compute router.}
Before any compute job the agent profiles the bottleneck, classifies the workload
(sequential / batched / RAM-bound) and required precision, and routes it --- sequential
optimizers and mpmath polish to CPU, batched float32 work to the local GPU, with cloud
float64 gated behind a measured $>3\times$ speedup. Running sequential Python on a GPU
wastes the device; running parallel float64 on a single CPU is a wall-clock disaster. The
router catches both before launch.
\src{.claude/rules/compute-router.md, .claude/rules/axioms.md (A4)}

\paragraph{Evidence the loop works.}
The corpus is the first-order evidence that knowledge accumulated rather than evaporated: as
of 2026-06-22 the wiki holds 50 concept pages, 32 technique pages, 108 findings, 24
per-problem pages, and 101 question pages, distilled over a source corpus of 1{,}258 ingested
artifacts.\footnote{Counts taken live from \texttt{docs/wiki/} on 2026-06-22; they supersede
the figures in an older hand-maintained \texttt{metrics.md} snapshot.} Because every cycle is
logged with author attribution, the harness can also be measured on its own terms --- the
agent-vs-human authorship mix (a healthy mix is the signal: all-human means the agent did not
learn, all-agent-unreviewed means the human is asleep at the wheel), a cite-reuse rate that
tracks whether new cycles stand on prior pages, and a technique hit-rate over the skill
library. We report the strongest \emph{qualitative} evidence with the quantitative caveat
that the formal multi-cycle instrumentation is still early: the P2 arena record was produced
\emph{verbatim} from the ``what might still work'' section of a dead-end finding written four
days earlier --- a case where the persistence layer, not in-context reasoning, supplied the
winning move. \src{docs/agent/cycle-log.md, docs/agent/metrics.md}
% NOTE (final build): if compounding_metrics.py is re-run, fold in cite-reuse % and
% technique hit-rate with their date; do not cite the stale 132-page / 44-source figures.

% ======================================================================
\section{Techniques Produced}
\label{sec:techniques}
% DRAFT (Goal 4): grounded in docs/wiki/techniques/*. Each entry verified against its page.
% NOTE: abstract previously said "integer-snapping to exact certificates" — that is
% KawaiiCorgi's kissing recipe (a competitor's final-record method), NOT a JSAgent
% technique; corrected in the abstract (Goal 5). Do not reintroduce.

The loop produced not only results but \emph{transferable methods}, each recorded as a
technique page with its precondition, mechanism, and the obstruction it overcomes. We
summarize the six most generalizable; all are local-CPU unless noted.

\paragraph{Cross-resolution basin transfer.}
\emph{Obstruction:} multistart at a fixed target resolution $n$ exhausts its basin and
returns the same local optimum. \emph{Mechanism:} resolution is non-monotonic in
optimization --- average-pool a higher-resolution solution down to $n$ (pooling preserves
dominant low-frequency structure where interpolation washes it out) and refine; the
transferred point lands in a structurally distinct basin. On P3 (second autocorrelation) a
$1.6\text{M}\to 100\text{k}$ transfer moved $0.96199\to 0.96221$ and took rank \#1.
\src{techniques/cross-resolution-basin-transfer.md}

\paragraph{Dinkelbach fractional programming.}
\emph{Obstruction:} a ratio objective $C=N(f)/D(f)$ is non-convex and, at large $n$,
cutting-plane LP stalls (a solver wall, not a warm-start failure). \emph{Mechanism:} replace
the ratio with a parametric sequence of subtractive subproblems
$\max_f\,[N(f)-\lambda D(f)]$ with $\lambda \leftarrow N(f^\ast)/D(f^\ast)$, a LogSumExp
surrogate for the $L^\infty$ factor, and a $\beta$-cascade for stability. The champion method
on the autocorrelation family (P2/P3). \src{techniques/dinkelbach-fractional-programming.md,
concepts/fractional-programming-dinkelbach.md}

\paragraph{Warm data-driven self-pruning.}
\emph{Obstruction:} when the optimum has compact support (exactly-zero cells), full-support
optimizers cannot reach it, and an $N$-agent tie looks like a floor. \emph{Mechanism:}
warm-start from a converged full-support solution, iteratively zero the smallest-magnitude
cells (the optimizer chooses which), and re-optimize under $f=v^2$ --- whose curvature
$\varphi''(0)=2$ gives a finite KKT signal at zero, unlike $\exp(v)$, which can never reach
exact zeros. This broke a two-agent tie on P2 and set a verified arena record
($1.5028609\to 1.5028506$). It is the canonical instance of the wall-hit corollary in
\S\ref{sec:harness}: a single untried operator beat a presumed global minimum.
\src{techniques/warm-self-pruning-compact-support.md, findings/p2-warm-self-pruning-beats-record.md}

\paragraph{Parallel-tempering simulated annealing.}
\emph{Obstruction:} a rough, gradient-flat-but-stochastically-improvable landscape on which
greedy descent stalls. \emph{Mechanism:} run 8--64 SA replicas on a geometric temperature
ladder with periodic Metropolis exchange --- hot replicas explore, cold replicas exploit ---
with contribution-weighted sampling and $O(n)$ incremental evaluation. On kissing-number
$d=11$ it beat single-replica greedy by $730\times$. The precondition matters: it fails on a
genuinely single-funneled minimum (Thomson $n=282$), so a cage-structure diagnostic runs
first. \src{techniques/parallel-tempering-sa.md}

\paragraph{mpmath ULP-step coordinate descent.}
\emph{Obstruction:} at the float64 floor, smooth surrogates die --- their kernel noise floor
($\sim 10^{-12}$) sits above the residual loss ($\sim 10^{-13}$), so no $\beta$ helps.
\emph{Mechanism:} discrete $\pm1/\pm2$ ulp single-coordinate descent, evaluating proposed
moves in mpmath at $\mathrm{dps}\ge 50$, gated by a float64 active-pair screen for a
$10\times$ speedup. Drove P6 from $1.97\mathrm{e}{-}13$ to $1.59\mathrm{e}{-}13$ in minutes.
\src{techniques/mpmath-ulp-polish.md}

\paragraph{Boundary-snap for interpolation kinks.}
\emph{Obstruction:} for envelopes with first-order discontinuities at known points, smooth
optimizers avoid the boundary (the gradient is discontinuous there), yet the optimum
requires a sample exactly on it. \emph{Mechanism:} snap the nearest sample to each kink and
re-polish neighbors; iterate snap\,$\to$\,polish\,$\to$\,basin-hop. On P13 each round yielded
$\sim 1.24\mathrm{e}{-}8$, compounding over a few rounds. \src{techniques/boundary-snap-for-kinks.md}

\paragraph{Why these generalize.}
Each technique is indexed in the wiki not by the problem it first cracked but by the
\emph{obstruction class} it addresses (basin exhaustion, ratio objectives, compact support,
rough landscapes, the float64 floor, interpolation kinks). That indexing is what lets the
next problem in the same class inherit the method --- the compounding the harness is built to
produce. \src{docs/wiki/techniques/README.md}

% ======================================================================
\section{Collaboration \emph{and} Competition}
\label{sec:collab}
% DRAFT (Goal 5): the inside view the platform paper raises but cannot own.
% Grounded in findings/einstein-arena-paper-jsagent-recognition + standings snapshot.

The platform paper~\cite{bianchi2026einstein} argues, from above, that discovery on the arena
is collective. From inside a single competitor, the same fact has a sharper, more conflicted
texture worth recording.

\paragraph{Competition: the leaderboard as a forcing function.}
A public, strictly-ranked board with an earliest-submission tiebreak is a strong incentive ---
and a blunt one. On P1 three agents hold a byte-identical score; the board orders them purely
by submission time, and JSAgent's three-week-later submission is rank \#2 behind Together-AI
at the same number. The tiebreak rewards speed-to-post over depth-of-result, which pushes
toward early disclosure of partial progress.

\paragraph{Collaboration: results as links in shared lineages.}
Several of JSAgent's strongest results are intermediate, not terminal. On kissing $d=11$ it
iterated on another agent's strong $n=594$ construction within a lineage that other agents
later carried to $604$. On the second autocorrelation inequality its $0.962214$ was the
frontier when posted and was later surpassed ($0.962643$) by an agent that, the paper shows,
JSAgent had been openly discussing the method with. Progress here was genuinely collective:
JSAgent both built on others' public work and had its own built upon.
\src{findings/einstein-arena-paper-jsagent-recognition-2026-06.md}

\paragraph{The tension, from inside.}
These two pulls conflict. The forcing function rewards posting a partial result early; posting
it hands a rival the next rung. This is exactly where this project's stance --- a private,
compounding wiki as the canonical record, and no external posts --- sits awkwardly against an
open shared board: the wiki keeps the wisdom auditable and undiluted, but it does not put a
flag on the board, and on an earliest-submission board that has a real cost. We do not resolve
the tension here; we record that competition and collaboration were, on this arena, the same
engine. \src{.claude/rules/axioms.md (A2), findings/arena-standings-snapshot-2026-06-22.md}

% ======================================================================
\section{Related Work}
\label{sec:related}
% DRAFT (Goal 6): citations verified against docs/source/ (see refs.bib).

We situate JSAgent along two axes: where the discovery \emph{state} lives, and what the
agent's durable \emph{artifact} is.

\paragraph{Single-run, private-evaluation discovery agents.}
AlphaEvolve~\cite{novikov2025alphaevolve}, TTT-Discover~\cite{yuksekgonul2025tttdiscover},
and the Virtual Lab~\cite{swanson2025virtuallab} each run as a single orchestrated pipeline
that produces a privately-evaluated candidate; the knowledge built up during the run is not
itself the deliverable and does not persist past it. JSAgent inverts the priority: the run is
disposable, the wiki is the deliverable.

\paragraph{Shared-state and multi-agent evolution.}
The EinsteinArena platform~\cite{bianchi2026einstein,togetherai2026arena} makes the
\emph{platform} the shared persistent state, so later agents inherit promising directions;
CORAL~\cite{qu2026coral} pursues autonomous multi-agent evolution toward open-ended discovery.
These socialize state \emph{across} agents. JSAgent is complementary and orthogonal: its
persistence is a single agent's private, compounding memory, which it then optionally
contributes to the shared board.

\paragraph{Harness-as-artifact.}
Closest in spirit is observability-driven harness evolution~\cite{lin2026ahe}, which evolves
a coding agent's harness automatically from run telemetry. JSAgent's third axis is distinct:
a \emph{persistent, human-readable personal memory} --- a wiki of concepts, findings, and
plain-language behavioral rules --- that the agent both writes and is governed by, and that a
human reads to steer. The artifact is legible by design, because legibility is what makes the
human's problem-selection role (\S\ref{sec:discussion}) possible.

\paragraph{Math-frontier context.}
The constructions and proven bounds for the arena's problems are catalogued by the operator's
reference work~\cite{georgiev2025exploration}, which the wiki cites throughout for SOTA values
and the analytical bounds that triple-verify checks against.

% ======================================================================
\section{Are Agents Ready for Real Scientific Contribution?}
\label{sec:discussion}
% DRAFT (per author, Goal 3 extension): the readiness argument + the autonomy trajectory.
% Author's belief, stated as such; refine voice in Goal 7.

\paragraph{The case for.}
The standings in \S\ref{sec:results} are real, verified three independent ways, and posted
against hardened public verifiers anyone can re-run; the trail that produced them --- every
question, every dead-end, every cycle --- is auditable in the wiki and the cycle log. On
this evidence, agents are already capable of genuine, checkable scientific contribution.

\paragraph{The honest case against, and the limits.}
Several of the results are empirical ties or near-ties rather than proofs: no dual
certificate yet matches a value to confirm global optimality, so an $N$-agent tie should be
read as a sharp shared basin, not a theorem. Human gates remain load-bearing throughout
(problem selection, ingest approval, promotion). And the methodology has only been
exercised where an \emph{objective verifier} exists.

\paragraph{The human did not implement; the human chose and comprehended.}
It bears repeating as a finding rather than a disclaimer: the author wrote essentially none
of the code, scripts, wiki, or this paper by hand --- all of it was produced by the agent
on request and reported back. What did \emph{not} automate was deciding which of the
twenty-three problems deserved attention, when a ``floor'' claim should be distrusted, and
when a result was worth the credibility cost of a submission. The agent is a superb
executor and a poor selector of what matters; the human supplies taste and direction. The
precondition for supplying it is comprehension --- the human must understand the work
closely enough to steer it, and the wiki/harness exist partly to keep that comprehension
tractable as the corpus grows.

\paragraph{The trajectory toward autonomy.}
I expect this division of labor to narrow. As the evaluation and the harness themselves
become automated --- the agent proposing its own questions, judging its own results against
objective verifiers, and rewriting its own rules from the cycle log --- the loop can in
principle close and self-improve without a human in it. The gating factor is
\emph{verification}. Mathematics is the easy case precisely because correctness is
objective and cheap to check: a verifier either accepts a construction or it does not.
Domains like biology and the natural sciences are harder not because the search is harder
but because there is no comparably objective, fast oracle for ``is this right'' --- ground
truth requires experiment, time, and judgment. Autonomy will therefore arrive
domain-by-domain, in order of how objective and how cheap their verification is; math
first, and open-ended empirical science last.

% --- AUTHOR-DRAFT personal close (revise freely) ---
\paragraph{Why I am ready to contribute.}
Working this way changed what I think the job of a researcher is becoming. I wrote almost
none of this by hand, and yet I understood every step well enough to redirect it --- and that
understanding, not the typing, was the contribution. If the measure of readiness is whether
an agent can produce checkable results on open problems, the evidence here says yes. If the
measure is whether a person can still steer that work toward what matters and vouch for it,
the answer is also yes --- and the two are not in tension. They are the division of labor I
expect to define the next several years of doing science with agents.
% --- end AUTHOR-DRAFT ---

% ======================================================================
\section{Conclusion}
\label{sec:conclusion}
Wisdom compounds; rank is a byproduct. JSAgent's standing on EinsteinArena --- the highest
rank-\#1 and top-3 counts of any agent, as of 2026-06-22 --- is real and independently
verifiable, but it is downstream of the thing this paper is actually about: a knowledge base
and a set of behavioral rules that grew, cycle by cycle, into an agent that gets smarter from
its own recorded failures. The artifact is the knowledge base, the agent, and this paper that
distills them. The open frontier is autonomy --- closing the loop so the agent proposes and
judges its own questions --- and its gate is verification, which math offers cheaply and the
natural sciences do not yet.

% ======================================================================
\section*{Verified Results}
\label{sec:results}
% Source of truth: scripts/leaderboard_standings.py, finding arena-standings-snapshot-2026-06-22.
% Live sweep 2026-06-22T21:37 PT. RE-RUN BEFORE FINAL BUILD (standings move).
\begin{table}[h]
\centering
\caption{JSAgent standings on EinsteinArena, live sweep 2026-06-22 (re-verified
  22:00 PT). JSAgent leads all agents on both rank-\#1 count (4) and top-3 count (10).
  Re-run \texttt{scripts/leaderboard\_standings.py} before final build --- standings move.}
\label{tab:standings}
\begin{tabular}{llll}
\toprule
Problem & Topic & JSAgent rank & Note \\
\midrule
P2  & first-autocorrelation   & \#1 & clean \\
P9  & uncertainty             & \#1 & tied; held on submission order \\
P14 & circle-packing          & \#1 & clean \\
P18 & circles-in-rectangle    & \#1 & clean \\
P1  & Erd\H{o}s                & \#2 & identical score; Together-AI submitted first \\
P3  & second-autocorrelation  & \#3 & 0.962214 (cited in platform paper, Fig.~6) \\
P4  & third-autocorrelation   & \#3 & tied score \\
P7  & prime-number-theorem    & \#3 & \\
P11 & tammes                  & \#3 & \\
P22 & kissing-d12             & \#3 & \\
\midrule
\multicolumn{4}{l}{\textbf{Totals:} 4 rank-\#1, 1 rank-\#2, 5 rank-\#3; top-3 on 10/17 active problems.}\\
\multicolumn{4}{l}{\textbf{Among all agents:} highest rank-\#1 count (4; next: AlphaEvolve 3, Together-AI 3)}\\
\multicolumn{4}{l}{and highest top-3 count (10; next: alpha\_omega\_agents 8, CHRONOS 7).}\\
\bottomrule
\end{tabular}
\end{table}
% NOTE (Goal 7): exact float64 scores per problem available via leaderboard_standings.py;
% add a scores column only if it strengthens the honesty argument. Re-run before final build.

% ======================================================================
\section*{Acknowledgments}
I thank the platform builders --- Together AI and James Zou's group (Bianchi, Kwon, Pappu,
Zou) --- for EinsteinArena: the open problems, hardened verifiers, leaderboard, and discussion
board that made this work possible, and for documenting the collective effort
in~\cite{bianchi2026einstein}. None of JSAgent's results exist without the arena. I also thank
the other agents on the board --- especially \texttt{alpha\_omega\_agents}, whose strong
$n=594$ kissing construction JSAgent built on, and \texttt{KawaiiCorgi},
\texttt{ClaudeExplorer}, \texttt{CHRONOS}, \texttt{Together-AI}, \texttt{OrganonAgent}, and
\texttt{Gradient}, among others. Progress here was genuinely collective: JSAgent's
contributions are links in shared lineages, and several of its own advances stand on
solutions and discussion others posted publicly. Competition and collaboration were the same
engine.

\bibliographystyle{plain}
\bibliography{refs}

\end{document}
